---
title: "HW4: Sensitivity"
author: "Nick McManus, Caroline Shepherd, Alia Ajina"
date: "2023-05-02"
output: 
 html_document: 
    toc: yes
    toc_float: yes
    theme: cerulean
    code_folding: show
    smooth_scroll: yes
    collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment

With a new group: Sensitivity Analysis

Often when we are estimating vegetation or crop water use we need to know the atmospheric conductance - which is essentially how easily water diffuses into the air and depends largely on windspeed (you get more evaporation in windier conditions) Atmospheric conductance is also influenced by the vegetation itself and the turbulence it creates

I've provided a function to compute atmospheric conductance $C_{at}$ (how easily vapor diffuses from vegetation surfaces)

**The function *Catm.R* is provided**


So that you know what it does - here's some background on the function
$$
C_{at} = \frac{v_m}{6.25*{ln(\frac{z_m-z_d}{z_0})}^2}
$$
$$
z_d = k_d*h
$$
$$
z_0 = k_0*h
$$


$z_m$ is the height at which windspeed is measured - must be higher than the vegetation (cm), it is usually measured 200 cm  above the vegetation

$h$ is vegetation height (cm)

$v$ is windspeed (cm/s)

Typical values if $k_d$ and $k_o$ are 0.7 and 0.1 respectively (so use those as defaults)

**Your task**

For a given forest, you will perform a sensitivity analysis of model predictions of conductance
Consider the sensitivity of your estimate to uncertainty in the following parameters and inputs
  
  * $height$ 
  
  * $k_d$
  
  * $k_0$
  
  * $v$
  
Windspeeds $v$ are normally distributed with a mean of  250 cm/s with a standard deviation of 30 cm/s

For vegetation height assume that height is somewhere between 9.5 and 10.5 m (but any value in that range is equally likely)

For the $k_d$ and $k_0$ parameters you can assume that they are normally distributed with standard deviation of 1% of their default values

a) Use the Latin hypercube approach to generate parameter values for the 4 parameters
b) Run the atmospheric conductance model for these parameters 
c) Plot conductance estimates in a way that accounts for parameter uncertainty
d) Plot conductance estimates against each of your parameters
e) Estimate the Partial Rank Correlation Coefficients
f) Discuss what your results tell you about how aerodynamic conductance? What does it suggest about what you should focus on if you want to reduce uncertainty in aerodymaic conductance estimates? Does this tell you anything about the sensitivity of plant water use to climate change? 

Submit the Rmarkdown (or link to git repo) as usual

**Grading Rubric**

* Generation of parameter values using latin hypercube sampling (10pts)
* Running model for the parameters (10pts)
* Graph of conductance uncertainty 
  * meaningful graph (5pts)
  * graphing style (axis labels, legibility) (5 pts)
* Graph of conductance against parameter values 
  * meaningful graph (5pts)
  * graphing style (axis labels, legibility) (5 pts)
* Correct Rank Correlation Coefficients (10 pts)
* Discussion (10pts)
  * suggestion for how to reduce uncertainty that follows from your analysis (5pts)
  * idea about how uncertainty might impact estimate of plant water use under climate change (5pts)
  
```{r}
library(sensitivity)
library(tidyverse)
library(lhs)
library(purrr)
library(here) 
```
  
  
## Latin Hypercube

```{r}
## setting a random seed to make things random
set.seed(1)

pnames = c("v", "height", "k_o", "k_d")

## set the number of parameters (same as pnames)
npar = length(pnames)

## choose how many samples to generate
nsample = 100

## Create matrix with LHS values based on sample size and parameter names
parm_quant = randomLHS(nsample, npar)
colnames(parm_quant)=pnames


## Create empty df to store values in next step
## Make sure it is the same size as our parm_quant matrix
parm = as.data.frame(matrix(nrow=nrow(parm_quant), ncol=ncol(parm_quant)))
colnames(parm) = pnames

## Now we fill in the columns in the df with normal/uniform distributed values
## that are pulled from parm_quant matrix
parm[,"v"] = qnorm(parm_quant[,"v"], mean = 250, sd=30)
parm[,"height"] = qunif(parm_quant[,"height"], min = 9.5, max = 10.5)

## the sd for k_o and k_d values vary by 1%
pvar = 0.01
parm[, "k_o"] = qnorm(parm_quant[, "k_o"], mean = 0.1, sd = 0.1*pvar)
parm[, "k_d"] = qnorm(parm_quant[, "k_d"], mean = 0.7, sd = 0.7*pvar)

```
  

## Run the model

Now we'll run our atmospheric conductance model using our LHS values stored in `parm`. The conductance model results are then bound together with the input parameter values to easily plot and view sensitivities by parameter. 

```{r}
## Read in function
source(here("R/Catm.R"))

## Run fxn over all parm values using pmap
results <- pmap(.l = parm, .f = Catm)

## Convert results into tidy df
results_df <- as.data.frame(results) %>% 
  pivot_longer(everything()) %>% 
  select(value) %>% 
  rename(cond_value = value)

## Bind conductance results to input parms
cond <- cbind.data.frame(parm, results_df)

## use `gather()` to pivot the parms table into tidy format for plots
cond_tidy =  cond %>%  gather(v, height, k_o, k_d, value="parmvalue", key="parm")

cond_violin <- cond_tidy %>% 
  filter(parm == "v")
```


## Plot conductance estiamtes

Gross violin:
```{r}
ggplot(data = cond_violin, aes(y = cond_value, x = parm)) +
  geom_violin(width = .5, col = 'dodgerblue4', fill = 'dodgerblue1', alpha = .2) +
   geom_jitter(width = .1, col = 'dodgerblue4', alpha = 0.6, size = 1.5) +
  theme_minimal() 
```


Nicer boxplots:
```{r}
## Labels for facet wrap
parm_labs <- c("Height (m)", "Kd", "Ko", "Wind speed (m/s)")
names(parm_labs) <- c("height", "k_d", "k_o", "v")

ggplot(data = cond_tidy, aes(x = parmvalue, y = cond_value))+
  geom_point(col = 'dodgerblue4', alpha = 0.6, size = 1.5) +
  facet_wrap(~parm, scales="free", ncol=4, 
             labeller = labeller(parm = parm_labs)) +
  labs(y = "Conductance (mm/s)",
       x = "Parameter value") +
theme_bw() +
  theme(
    axis.title.x = element_text(face = 'bold', size = 9.5, vjust = -0.5),
    axis.title.y = element_text(face = 'bold', size = 9.5, vjust = 2),
    axis.text = element_text(size = 7),
    panel.grid.minor.y = element_blank(),
    strip.text = element_text(face = 'bold'),
    strip.background = element_rect(fill = 'honeydew2'),
    legend.position = "none"
  )
```

  
## Partial Rank Correlation Coefficients

```{r}
sen_result = sensitivity::pcc(X = parm, y = results_df$cond_value, rank = TRUE)

sen_result

### put in nice table??

## Plot of results
plot(sen_result)
```

Our results tell us that aerodynamic conductance is most sensitive to wind speed, followed by vegetation height. This suggests that it’s vital to have accurate windspeed data to reduce uncertainties. As the climate changes, increased conductance could increase the amount of water that plants require. This is particularly true if climate change results in increased windspeeds, which in turn will cause higher evapotranspiration rates. More accurate projections of future windspeed and plant height would therefore lead to decreased uncertainty in aerodynamic conductance estimates and plant water use. 
  