---
title: "HW4: Sensitivity"
author: "Nick McManus, Caroline Shepherd, Alia Ajina"
date: "2023-05-02"
output: 
 html_document: 
    toc: yes
    toc_float: yes
    theme: cerulean
    code_folding: show
    smooth_scroll: yes
    collapsed: yes
---

# Assignment 4: Sensitivity


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE, warning = FALSE)

library(sensitivity)
library(tidyverse)
library(lhs)
library(purrr)
library(here)
library(ggpubr)
```
  
  
## Latin Hypercube

```{r}
## setting a random seed to make things random
set.seed(1)

pnames = c("v", "height", "k_o", "k_d")

## set the number of parameters (same as pnames)
npar = length(pnames)

## choose how many samples to generate
nsample = 100

## Create matrix with LHS values based on sample size and parameter names
parm_quant = randomLHS(nsample, npar)
colnames(parm_quant)=pnames


## Create empty df to store values in next step
## Make sure it is the same size as our parm_quant matrix
parm = as.data.frame(matrix(nrow=nrow(parm_quant), ncol=ncol(parm_quant)))
colnames(parm) = pnames

## Now we fill in the columns in the df with normal/uniform distributed values
## that are pulled from parm_quant matrix
parm[,"v"] = qnorm(parm_quant[,"v"], mean = 250, sd=30)
parm[,"height"] = qunif(parm_quant[,"height"], min = 9.5, max = 10.5)

## the sd for k_o and k_d values vary by 1%
pvar = 0.01
parm[, "k_o"] = qnorm(parm_quant[, "k_o"], mean = 0.1, sd = 0.1*pvar)
parm[, "k_d"] = qnorm(parm_quant[, "k_d"], mean = 0.7, sd = 0.7*pvar)

```
  

## Run the model

Now we'll run our atmospheric conductance model using our LHS values stored in `parm`. The conductance model results are then bound together with the input parameter values to easily plot and view sensitivities by parameter. 

```{r}
## Read in function
source(here("R/Catm.R"))

## Run fxn over all parm values using pmap
results <- pmap(.l = parm, .f = Catm)

## Convert results into tidy df
results_df <- as.data.frame(results) %>% 
  pivot_longer(everything()) %>% 
  select(value) %>% 
  rename(cond_value = value)

## Bind conductance results to input parms
cond <- cbind.data.frame(parm, results_df)

## use `gather()` to pivot the parms table into tidy format for plots
cond_tidy =  cond %>%  gather(v, height, k_o, k_d, value="parmvalue", key="parm")

cond_violin <- cond_tidy %>% 
  filter(parm == "v")
```


## Plot conductance estiamtes

```{r}
# Density plots to display parameter uncertainty input values
p1<-ggplot(data = cond, aes(x = height)) +
  geom_density()

p2<-ggplot(data=cond, aes(x=v)) +
  geom_density()

p3<-ggplot(data=cond, aes(x=k_o)) +
  geom_density()

p4<-ggplot(data=cond, aes(x=k_d)) +
  geom_density()

ggarrange(p1, p2, p3, p4)
```


```{r}
## Labels for facet wrap
parm_labs <- c("Height (m)", "Kd", "Ko", "Wind speed (m/s)")
names(parm_labs) <- c("height", "k_d", "k_o", "v")

ggplot(data = cond_tidy, aes(x = parmvalue, y = cond_value))+
  geom_point(col = 'dodgerblue4', alpha = 0.6, size = 1.5) +
  facet_wrap(~parm, scales="free", ncol=4, 
             labeller = labeller(parm = parm_labs)) +
  labs(y = "Conductance (mm/s)",
       x = "Parameter value") +
theme_bw() +
  theme(
    axis.title.x = element_text(face = 'bold', size = 9.5, vjust = -0.5),
    axis.title.y = element_text(face = 'bold', size = 9.5, vjust = 2),
    axis.text = element_text(size = 7),
    panel.grid.minor.y = element_blank(),
    strip.text = element_text(face = 'bold'),
    strip.background = element_rect(fill = 'honeydew2'),
    legend.position = "none"
  )
```

  
## Partial Rank Correlation Coefficients

```{r}
sen_result = sensitivity::pcc(X = parm, y = results_df$cond_value, rank = TRUE)

sen_result

## Plot of results
plot(sen_result)
```

Our results tell us that aerodynamic conductance is most sensitive to wind speed, followed by vegetation height. This suggests that it’s vital to have accurate windspeed data to reduce uncertainties. As the climate changes, increased conductance could increase the amount of water that plants require. This is particularly true if climate change results in increased windspeeds, which in turn will cause higher evapotranspiration rates. More accurate projections of future windspeed and plant height would therefore lead to decreased uncertainty in aerodynamic conductance estimates and plant water use. 
  