---
title: "HW3: Almond Profit and Informal Sensitivity Analysis"
author: "Katheryn Moya & Nick McManus"
date: "4/27/2023"
output: 
 html_document: 
    toc: yes
    toc_float: yes
    theme: cerulean
    code_folding: show
    smooth_scroll: yes
    collapsed: yes
---

```{r setup, include=T, warning=F, message=F}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)

library(tidyverse)
library(here)
library(purrr)
```


# Read in and wrangle climate data

First we'll read in our `clim.txt` data file, which provides daily precipitation as well as minimum and maximum temperatures. Next, we'll create a data subset with the total precipitation and absolute minimum and maximum temperatures for each month. Then, we'll form two subsets for use in the almond yield function: one with minimum February temperatures and another with the total January precipitation.
```{r}
### Read in data
clim_df <- read_table("clim.txt") %>% 
  ## remove quotes in names
  janitor::clean_names()


### Find the absolute tmin and tmax and summed precip
clim_months <- clim_df %>% 
  group_by(month, year) %>% 
  summarize(tmax_max = max(tmax_c),
            tmin_min = min(tmin_c),
            precip_sum = sum(precip))


### Filter for Feb Tmin
tmin_feb <- clim_months %>% 
  filter(month ==2) %>% 
  ungroup() %>% 
  select(year, tmin_min)

### Filter for Jan precip
precip_jan <- clim_months %>% 
  filter(month == 1) %>% 
  ungroup() %>% 
  select(year, precip_sum)
```



# Sensitivity Analysis

We'll perform the sensitivity analysis using outputs from the almond yield function. This function outputs the profit based on yield anomaly for each year within the time series. 

```{r}
### Read in function
source(here("R/almond_yield_profit.R"))
```


##  Generate samples

For this analysis, we're testing for the uncertainties surrounding the `tcoeff1` and `pcoeff2` parameters -- which affect almond yield -- as well as the price per ton/acre (`price_ton`) for almonds. The informal sensitivity analysis for the `tcoeff1` parameter includes the following assumptions:

* Default value of the parameter is -0.015
* Uncertainty of $\pm$20% for parameter value
* Uniform distribution since we do not know anything about variation

For `pcoeff2`, the following assumptions were used to generate a sample of values:
* Default value of parameter is 0.0043
* Uncertainty of $\pm$20% for parameter value
* Uniform distribution since we do not know anything about variation

For the price per ton/acre of almonds, the following assumptions were made to randomly generate samples:
* Default (mean) price is $3520/ton/acre
* Due to market changes, one standard deviation in price is $200/ton/acre
* Changes in price follow normal distribution

```{r}
## Number of samples
nsamples = 300
## Percent deviation from baseline
dev = 0.2


### Generate tcoeff1 samples
tcoeff1_base = -0.015
tcoeff1 <- runif(min = tcoeff1_base + tcoeff1_base*dev,
                        max = tcoeff1_base - tcoeff1_base*dev,
                        n = nsamples)

### Generate pcoeff2 sample
pcoeff2 <- 0.0043
pcoeff2 <- runif(min = pcoeff2 - pcoeff2*dev,
                 max = pcoeff2 + pcoeff2*dev,
                 n = nsamples)

### Generate price samples
price_ton <- rnorm(mean = 3520, sd = 100, n = nsamples)



### bind samples together in one df
params <- cbind.data.frame(tcoeff1, pcoeff2, price_ton)
```


## Run function with `purrr::pmap()`

Using `pmap()`, we can easily run our almond yield function across our entire sample of parameter values (300 times!). The resulting df `profit` contains the yearly profit generated for each sample value of `tcoeff1`, `pcoeff2`, and `price_ton`. This results in 300 profit values per year over the 21 year time period.
```{r}
### run with pmap
results <- params %>% pmap(almond_yield_profit, 
                           tmin = tmin_feb$tmin_min,
                           precip = precip_jan$precip_sum,
                           year = tmin_feb$year)

### View head of results and make sure length = 300
# results[[1]]
# length(results)

### Extract results from list into df
profit <- map_df(results, `[`, c("year", "profit")) 


# profit_subset <- profit %>% 
#   filter(!year %in% c(1995, 1997, 2001, 2005, 2008, 2010))

### Exploratory stuff to see which parameters have biggest effect

# ### Bind results to the input parameter values
# profit <- cbind.data.frame(profit, params)
#
#
# p1 = ggplot(yield_profit, aes(x = tcoeff1, y = mean_profit, col=pcoeff2))+
#   geom_point(cex=2)+
#   labs(y="Profit (USD/TON/ACRE)", x="tcoeff1")
# p2 = ggplot(yield_profit, aes(x = tcoeff2, y = mean_profit, col=pcoeff2))+
#   geom_point(cex=2)+
#   labs(y="Profit (USD/TON/ACRE)", x="tcoeff2")
# p3 = ggplot(yield_profit, aes(x = pcoeff1, y = mean_profit, col=tcoeff1))+
#   geom_point(cex=2)+
#   labs(y="Profit (USD/TON/ACRE)", x="pcoeff1")
# p4 = ggplot(yield_profit, aes(x = pcoeff2, y = mean_profit, col=tcoeff1))+
#   geom_point(cex=2)+
#   labs(y="Profit (USD/TON/ACRE)", x="pcoeff2")
# ggarrange(p1,p2,p3, p4)
```



## Visualize results

Finally, the profit for each year are visualized to informally view the uncertainty of almond profit based on three parameters.
```{r fig.width = 10, fig.height = 5}
ggplot(profit, aes(x = as.factor(year), y = profit, group = year)) +
  geom_boxplot(fill = 'dodgerblue2', color = 'dodgerblue4', 
               alpha = 0.6, width = 0.9) + 
  labs(y="Profit (USD/ton/acre)", x="Year") +
  # scale_y_log10()+
  theme_minimal() +
  theme(
    axis.title.x = element_text(face = 'bold', size = 11, vjust = -0.4),
    axis.title.y = element_text(face = 'bold', size = 11, vjust = 2),
    axis.text.x = element_text(size = 10)
  )
```

**Figure 1.** Distribution of annual profits (USD/ton/acre) for almond yield between 1989 and 2010 based on sampled parameters for precipitation, temperature, and price. Mean profits remain fairly consistent across all years with the largest variation in profits driven by an increase in total precipitation (mm).



