---
title: "HW3: Almond Profit and Informal Sensitivity Analysis"
author: "Katheryn Moya & Nick McManus"
date: "4/27/2023"
output: 
 html_document: 
    toc: yes
    toc_float: yes
    theme: cerulean
    code_folding: hide
    smooth_scroll: yes
    collapsed: yes
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(purrr)
library(ggpubr)
```


## Read in and wrangle climate data

First we'll read in our `clim.txt` data file, which provides daily precipitation as well as minimum and maximum temperatures. Next, we'll create a data subset with the total precipitation and absolute minimum and maximum temperatures for each month. Then, we'll form two subsets for use in the almond yield function: one with minimum February temperatures and another with the total January precipitation.
```{r}
### Read in data
clim_df <- read_table("clim.txt") %>% 
  ## remove quotes in names
  janitor::clean_names()


### Find the absolute tmin and tmax and summed precip
clim_months <- clim_df %>% 
  group_by(month, year) %>% 
  summarize(tmax_max = max(tmax_c),
            tmin_min = min(tmin_c),
            precip_sum = sum(precip))


### Filter for Feb Tmin
tmin_feb <- (clim_months %>% filter(month ==2))$tmin_min 

### Filter for Jan precip
jan_precip <- (clim_months %>% filter(month ==1))$precip_sum
```



# Sensitivity Analysis

We'll perform the sensitivity analysis using outputs from the almond yield function. This function outputs the yield anomaly and profit for each year within the time series. 

```{r}
### Read in function
source(here("R/almond_yield_profit.R"))
```


##  Generate samples

For this analysis, we're testing for the uncertainties surrounding the *tcoeff1* parameter -- which affects almond yield anomaly -- as well as the price for almonds. The informal sensitivity analysis for the *tcoeff1* parameter includes the following assumptions:

* Default value of the parameter is -0.015
* Uncertainty of  $\pm$20% for parameter
* Uniform distribution since we do not know anything about variation

For the price per ton per acre of almonds, the following assumptions were made to randomly generate samples:
* Default (mean) price is $3520/ton/acre
* Due to market changes, one standard deviation in price is $200/ton/acre
* Changes in price follow normal distribution

```{r}
nsamples = 300

### tcoeff1 sample
dev = 0.2
tcoeff1_base = -0.015

tcoeff1_sample <- runif(min = tcoeff1_base + tcoeff1_base*dev,
                        max = tcoeff1_base - tcoeff1_base*dev,
                        n = nsamples)

### price samples
price_sample <- rnorm(mean = 3520, sd = 200, n = nsamples)


### bind samples together in one df
parms <- cbind.data.frame(tcoeff1_sample, price_sample)
```


## Run function with `purrr::pmap`
```{r}
### run with pmap
```



## Visualize results
Here we will make the "meaningful" figures
